Moral Foundations of U.S. Political News Organizations

### 2019-01-01 pilot_manuscript.Rmd
More in Exp1 Method and chose sources!

@@

Political articles were scraped from the websites of four notable U.S. news sources. The sources were *The New York Times*, *National Public Radio (NPR)*, *Fox News*, and *Breitbart*. They were selected for their widespread recognition and the fact they are easily categorized (by the general public) according to perceived political lean. In general, *The New York Times* and *NPR* are perceived by many to have a liberal bias or lean. In contrast, Fox News and Breitbart are believed to have a conservative bias or lean. Political articles in particular were identified and subsequently scraped by including the specific URL directing to each source's political content in the *R* script. For example, rather than scrape from nytimes.com, which would return undesired results (non-political features, reviews, etc.), we instead included nytimes.com/section/politics so that more or less exclusively political content was obtained. All code for this manuscript can be found at OSF LINK, and the scripts are provided inline with this manuscript written with the *papaja* library [@Aust2017].

@@

Using the @Warriner2013 dictionary, the words making up each of the five foundations in the MFD were assigned their respective valence value. Once the MFD words' valences were added to the dataset, the article text was processed using the *tm* and *ngram* packages in *R* in order to render the text in lowercase, remove punctuation, and fix spacing issues [@CITE_tm]. The individual words were then reduced to their stems (i.e., *abused* was stemmed to *abus*). The same procedure was applied to the MFD words and the words in the @Warriner2013 dataset.

@@

The source article words were compiled into a dataset where they were matched up with their counterparts in the MFD along with their valence and a percentage of their occurrence. Therefore, for each article, the percentage of the number of *harm/care* words occurring in the articles were calculated, and this process was repeated for each of the foundations. This procedure created five percentages that were included as the dependent variable for the following analyses.

@@

For Experiment 2, the researchers hypothesized that news sources perceived as liberal will exhibit positive endorsements of the individualizing moral foundations (*harm/care* and *fairness/reciprocity*) in their articles reporting on the Kavanaugh confirmation hearing. News sources perceived as conservative are hypothesized to positively endorse all five foundations equally in their coverage of the Kavanaugh hearing. The reasearchers tested the hypothesis by analyzing the content scraped from news sources' web pages during the two weeks Kavanaugh's confirmation was prominent in the news. The content will be analyzed for valence and moral alignment under Moral Foundations Theory.

@@

Expected material stuff
- we are going to pick liberal and conservative sources from that thing document linked stuff
- list those here:
  Sources used by LIBERALS (according to the document thing):
  The New Yorker
  Slate
  The Daily Show (is it wise to use satirical news for this??)
  The Guardian
  Al Jazeera America
  NPR
  New York Times
  Buzzfeed
  PBS
  BBC
  Huffington Post
  Washington Post
  The Economist
  Politico
  MSNBC
  CNN
  NBC News
  CBS News
  Google News
  Bloomberg
  ABC News
  USA TODAY

  Sources used by CONSERVATIVES:
  Fox News
  Drudge Report
  Breitbart
  Rush Limbaugh Show
  The Blaze
  Sean Hannity
  Glenn Beck (does he still report "news"??)

- pick a specific date range we want to pull articles from 
- list that here:

>>>>

# Method
For Experiment 1, the researchers approached the methodology with the intention to answer a method question. That is, this portion of the current research was conducted in order to solidify the best method by which to analyze political news text under the Moral Foundations Theory framework. The researchers hypothesized the news sources genrally perceived as liberal leaning (*NPR* and *The New York Times*) would contain MFD words and valences indicating endorsements of the individualizing moral foundations (*harm/care* and *fairness/reciprocity*). Additionally, the researchers hypothesized the two sources generally perceived to be conservative leaning (*Fox News* and *Breitbart*) would feature MFD words and valences indicating equal endorsement of all five foundations.

@@

Political articles were scraped from the websites of four notable U.S. news sources. The sources were *The New York Times*, *National Public Radio (NPR)*, *Fox News*, and *Breitbart*. They were selected for their widespread recognition and the fact they are easily categorized (by the general public) according to perceived political lean. In general, *The New York Times* and *NPR* are perceived by many to have a liberal bias or lean. In contrast, Fox News and Breitbart are believed to have a conservative bias or lean [@Mitchell2014]. Political articles in particular were identified and subsequently scraped by including the specific URL directing to each source's political content in the *R* script. For example, rather than scrape from nytimes.com, which would return undesired results (non-political features, reviews, etc.), we instead included nytimes.com/section/politics so that more or less exclusively political content was obtained. All code for this manuscript can be found at OSF LINK, and the scripts are provided inline with this manuscript written with the *papaja* library [@Aust2017].

@@

The article text was processed using the *tm* and *ngram* packages in *R* in order to render the text in lowercase, remove punctuation, and fix spacing issues [@CITE_tm]. The individual words were then reduced to their stems (i.e., *abused* was stemmed to *abus*). The same procedure was applied to the MFD words and the words in the @Warriner2013 dataset. Using the @Warriner2013 dictionary, the words making up each of the five foundations in the MFD were assigned their respective valence value.

@@

The source article words were compiled into a dataset where they were matched up with their counterparts in the MFD along with their valence and a percentage of their occurrence. Therefore, for each article, the percentage of the number of *harm/care* words occurring in the articles were calculated, and this process was repeated for each of the foundations. This procedure created five percentages that were included as the dependent variable for the following analyses. It should be noted the valence scores for the MFD words were *z*-scored, so negative valence scores are possible.

@@

In contrast to Experiment 1, the reserachers approached Experiment 2 with the intention to confirm the method employed was valid for the analysis of the scaped text as well as for any inferences drawn from the analyses. For Experiment 2, the researchers hypothesized that news sources perceived as liberal will exhibit positive endorsements of the individualizing moral foundations (*harm/care* and *fairness/reciprocity*) in their articles reporting on the Kavanaugh confirmation hearing. News sources perceived as conservative are hypothesized to positively endorse all five foundations equally in their coverage of the Kavanaugh hearing. The reasearchers tested the hypothesis by analyzing the content scraped from news sources' web pages during the two weeks Kavanaugh's confirmation was prominent in the news. The content will be analyzed for valence and moral alignment under Moral Foundations Theory.

@@

Expected material stuff
- we are going to pick liberal and conservative sources from that thing document linked stuff
- list those here:
  Sources used by LIBERALS:
  MSNBC
  Slate
  Huffington Post
  NPR
  New York Times
  Politico

  Sources used by CONSERVATIVES:
  Fox News
  Drudge Report
  Breitbart
  Rush Limbaugh Show
  The Blaze
  Sean Hannity

  DROPPED:
  Glenn Beck (does he still report "news"??)
  The Daily Show (is it wise to use satirical news for this??) - Answer = Nope
  The Guardian (not American)
  The New Yorker
  Al Jazeera America
  Buzzfeed
  PBS
  BBC
  Washington Post
  The Economist
  CNN
  NBC News
  CBS News
  Google News
  Bloomberg
  ABC News
  USA TODAY
